---
title: "Machine Learning Final"
author: "Tim Lambert"
date: "5/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Approach

Using exercise data, we attempt to predict how a particular subject performed an exercise.  By downloading the data from the website, getting rid of blank or erroneous data, and separating it into a training and test set, training and cross-validation can be performed using the caret and RandomForest packages in R.

```{r getting_cleaning_data}
library(caret)
library(randomForest)

set.seed(10000)

training.data <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",
                          na.strings=c("NA","#DIV/0!", ""))
testing.data <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",
                         na.strings=c("NA","#DIV/0!", ""))
                       
#Delete columns with zero
training.data <- training.data[,colSums(is.na(training.data)) == 0]

#Number of NA values
sum(is.na(training.data))

#Delete admin data
training.data <- training.data[,-c(1:7)]

#Create training and test set for cross validation
index <- createDataPartition(training.data$classe, p = .75, list = FALSE)

#Create training and test set
training.data.train <- training.data[index,]
training.data.test <- training.data[-index,]
rm(index)


```

## Model One - Decision Tree

First, a decision tree model is applied to the training set and a cross validation is applied to test set.  Then, a confusion matrix is used to see how well the model did.

```{r dtm}
#Decision Tree Model
modfit1 <- train(data = training.data.train, classe ~ ., method = "rpart")

#Decision Tree Confusion Matrix
predict.modfit1 <- predict(modfit1, training.data.test)
confusionMatrix(training.data.test$classe, predict.modfit1)
```

The decision tree model's accuracy is just shy of 50% meaning the out of sample error is just above 50%.

##Random Forest Model

Next a Random Forest is applied to the training data set.  Just as with the decision tree model, cross-validation is applied to test set and a confusion matrix used to see how well the model performed.

```{r rfm}
#Random Forest Model
modfit2 <- randomForest(data = training.data.train, classe ~ .)

#Random Forest Confusion Matrix
predict.modfit2 <- predict(modfit2, training.data.test)
confusionMatrix(training.data.test$classe, predict.modfit2)
```

The random forest did much better with an accuracy of 99.59% meaning the out of sample error is .41%.

##20 observations

Use the random forest model to predict the 20 observations.

```{r test}
#Predict the 20 from test data set
predict(modfit2, testing.data)
```